---
title: "R Notebook"
output: html_notebook
---

#### 1. Setting the working directory
```{r}
setwd("/Users/busara/Documents/Personal Development/R/Titanic")
seed_object = 331
```

####2. Loading the required packages
```{r}
pkgs <- c("caret", 'caretEnsemble', "dplyr", "readr", "data.table", 
          "lubridate", "imputeMissings", "VIM","Hmisc","mice","ROSE")
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] # vector of missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}
invisible(lapply(pkgs,library,character.only=TRUE))
```


####3. Reading in the data
```{r}
train.dataset = read_csv("train.csv")
train.dataset$data = "train"
test.dataset = read_csv("test.csv")
test.dataset$data = "test"
```

####4. Combine the dataset so that we can clean the data at once
```{r}
titanic_dataset = bind_rows(train.dataset,test.dataset)
```

####5. Drop some variables that are not really important
```{r}
titanic_dataset<-titanic_dataset%>%
  dplyr::select(-Name,-Ticket,-Cabin)
```

####6. Describe the data
```{r}
Hmisc::describe(titanic_dataset)

missingness = data.frame(sapply(titanic_dataset,function(x) sum(is.na(x))))

#Age and Survived have the highest missingness so we need to impute.
```
####7. Convert the variables to the appropriate data types
```{r}
cats = c("Pclass","Sex","SibSp","Parch","Embarked")
#factor
titanic_dataset$Survived<-as.factor(titanic_dataset$Survived)
titanic_dataset$Pclass<-as.factor(titanic_dataset$Pclass)
titanic_dataset$Sex<-as.factor(titanic_dataset$Sex)
titanic_dataset$SibSp<-as.factor(titanic_dataset$SibSp)
titanic_dataset$Parch<-as.factor(titanic_dataset$Parch)
titanic_dataset$Embarked<-as.factor(titanic_dataset$Embarked)
```


####8. Imputation of missing values using mice
```{r}
titanic_imputation<- mice(titanic_dataset[, c("Age","Embarked","Fare")], method = "cart", seed = 201,m = 30)

titanic_imputation <- complete(titanic_imputation, sample(c(1:30), 1))
titanic_dataset2 <- cbind.data.frame(titanic_dataset[, c(-5,-8,-9)], titanic_imputation)

missingness2 = data.frame(sapply(titanic_dataset2,function(x) sum(is.na(x))))

```


####9. Generate some variables, and drop others
```{r}
## Generate a variable that indicates whether one had a sibling in the ship or not
titanic_dataset2$SibSp2 = ifelse(titanic_dataset2$SibSp ==0, "No","Yes")
titanic_dataset2$SibSp2<-as.factor(titanic_dataset2$SibSp2)
titanic_dataset2$SibSp<-NULL
## Generate a variable that indicates whether one had a parent or child in the ship or not
titanic_dataset2$Parch2 = ifelse(titanic_dataset2$Parch ==0, "No","Yes")
titanic_dataset2$Parch2<-as.factor(titanic_dataset2$Parch2)

titanic_dataset2$Parch<-NULL
```


####10. Split the data back into train and test
```{r}
train.dataset2 = titanic_dataset2[1:891,]
train.dataset2$data<-NULL
test.dataset2 = titanic_dataset2[892:1309,]
test.dataset2$data<-NULL
```

####10. Take care of imbalanced data
```{r}
train.dataset.rose <- ROSE(Survived ~ ., data = train.dataset2[,-1], seed = 567)$data
#train.dataset.rose$PassengerId<-train.dataset2$PassengerId
table(train.dataset.rose$Survived)
```

####11. Create the train control object
```{r}
myControl <- trainControl(
  method = "repeatedcv", number=10, repeats=7,
  classProbs = FALSE, 
  verboseIter = TRUE,
  savePredictions = TRUE
)

```

####12. Fit a ridge or lasso model
```{r}
train.dataset.rose_glmnet<-train.dataset.rose
set.seed(seed_object)
model_glmnet <- train(
  Survived ~ ., data = train.dataset.rose_glmnet,
  metric = "Accuracy",
  method = "glmnet",
  trControl = myControl,
  preProcess = c('zv', 'center', 'scale')
)

```

####13. Predict the values using test dataset, for the glmnet model
```{r}
test.dataset2.rose_glmnet<-test.dataset2
test.dataset2.rose_glmnet$Survived <- predict(model_glmnet, newdata = test.dataset2.rose_glmnet)


train.dataset.rose_glmnet$Survived2 <- predict(model_glmnet, newdata = train.dataset.rose_glmnet)
conf.matrix<-table(train.dataset.rose_glmnet$Survived,train.dataset.rose_glmnet$Survived2)
conf.matrix

accuracy = (sum(conf.matrix[1],conf.matrix[4])/sum(conf.matrix))
accuracy

titanic_submission<-test.dataset2.rose_glmnet%>%
  dplyr::select(PassengerId,Survived)
write.csv(titanic_submission,"titanic_submission_glmnet4.csv")
## training data accuracy on glmnet is 0.7800224

##The score is 0.76555

```
####14. XGBoost model
```{r}
#Need to tune on two parameters
train.dataset.rose_xgb = train.dataset.rose
set.seed(seed_object)
model_xgb <-train(
  Survived ~ ., data = train.dataset.rose_xgb,
  metric = "Accuracy",
  method = "xgbTree",
  trControl = myControl,
  preProcess = c('zv', 'center', 'scale', 'pca')
)
```

####15.Predict the values using test dataset, for the XGBoost model
```{r}

test.dataset2.rose_xgb<-test.dataset2
test.dataset2.rose_xgb$Survived <- predict(model_xgb, newdata = test.dataset2.rose_xgb)


train.dataset.rose_xgb$Survived2<- predict(model_xgb, newdata = train.dataset.rose_xgb)
conf.matrix<-table(train.dataset.rose_xgb$Survived,train.dataset.rose_xgb$Survived2)
conf.matrix

accuracy = (sum(conf.matrix[1],conf.matrix[4])/sum(conf.matrix))
accuracy
## training data accuracy on xgb is 0.9034792

titanic_submission<-test.dataset2.rose_xgb%>%
  dplyr::select(PassengerId,Survived)
write.csv(titanic_submission,"titanic_submission_xgb2.csv")

## The score is 0.73684
```

####16. Random forest model
```{r}
#Need to tune on two parameters
set.seed(seed_object)
model_rf <- train(
  Survived ~ ., data = train.dataset.rose,
  metric = "Accuracy",
  tuneLength = 10,
  method = "ranger",
  trControl = myControl,
  preProcess = c('zv', 'center', 'scale', 'pca')
)

```

```{r}
train.dataset.rose_rf<-train.dataset.rose
test.dataset2.rose_rf<-test.dataset2
test.dataset2.rose_rf$Survived <- predict(model_xgb, newdata = test.dataset2.rose_rf)


train.dataset.rose_rf$Survived2<- predict(model_xgb, newdata = train.dataset.rose_rf)
conf.matrix<-table(train.dataset.rose_rf$Survived,train.dataset.rose_rf$Survived2)
conf.matrix

accuracy = (sum(conf.matrix[1],conf.matrix[4])/sum(conf.matrix))
accuracy
## training data accuracy on rf is 0.9113356

titanic_submission<-test.dataset2.rose_rf%>%
  dplyr::select(PassengerId,Survived)
write.csv(titanic_submission,"titanic_submission_rf2.csv")
```
